---
title: "Chapter 14 Transformation"
author: "kwon"
date: "2018년 6월 29일"
output: html_document
---

# 1. Response Transformation

```{r}
initech = read.csv("initech.csv")

plot(salary~ years, data= initech, col="grey", pch=20, cex=1.5, main = "Salaries at Initech, by Seniority")
```
```{r}
initech_fit = lm(salary ~ years, data= initech)
summary(initech_fit)

plot(salary~years, data= initech, col="grey", pch=20, cex=1.5, main ="Salaries at Initech")
abline(initech_fit, col="darkorange", lwd=2)
```

```{r fig.width=10}
par(mfrow= c(1,2))

plot(fitted(initech_fit), resid(initech_fit), col="grey", pch=20, xlab="Fitted", ylab="Residuals", main= " Fitted versus Residuals")
abline(h=0, col="darkorange", lwd=2)

qqnorm(resid(initech_fit), main="Normal Q-Q Plot", col= "darkgrey")
qqline(resid(initech_fit), col ="dodgerblue", lwd=2)
```

## 1.1 Variance Stablizing Transformations
$$ \epsilon \sim N(0, \sigma^2) $$
Thus,
$$ Var[Y| X=x] = \sigma^2$$
However, here we see that the vairance is a function of the mean
$$ Var[Y| X=x] = h(E[Y|X=x])) $$
In order to corret for this, we would like to find some funciton of Y,
$$ Var[g(Y) | X =x ]= c $$
This is called a variance stabilizing transformation(VST). \
A common VST when we see increasing vairance in a fitted versus residuals plot is log(Y).
$$ log(Y_i) = \beta_0 + \beta_1x_i + \epsilon_i$$
$$ Y_i = exp(\beta_0 + \beta_1 x_i)exp(\epsilon_i)$$
```{r}
initech_fit_log = lm(log(salary)~years, data=initech)

plot(log(salary)~ years, data=initech, col = "grey", pch =20, cex = 1.5, main ="Salaries at Initech, By Seniority")
abline(initech_fit_log, col="darkorange", lwd=2)
```

```{r}
plot(salary~years, data= initech, col="grey", pch = 20, cex = 1.5, main = "Salaries at Initech, By Seniority")
curve(exp(initech_fit_log$coef[1] + initech_fit_log$coef[2]*x), from = 0, to = 30, add= TRUE, col ="darkorange", lwd=2)
```
```{r}
par(mfrow = c(1, 2))

plot(fitted(initech_fit_log), resid(initech_fit_log), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(initech_fit_log), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(initech_fit_log), col = "dodgerblue", lwd = 2)
```

```{r}
sqrt(mean(resid(initech_fit)^2))
sqrt(mean((initech$salary - exp(fitted(initech_fit_log))) ^ 2))
```

## 1.2 Box-Cox Transformations
$$ g_\lambda(y) =  \left\{
\begin{array}{lr}\displaystyle\frac{y^\lambda - 1}{\lambda} &  \lambda \neq 0\\
        & \\
       \log(y) &  \lambda = 0
     \end{array}
   \right.
$$
\[
  L(\lambda) = -\frac{n}{2}\log(RSS_\lambda / n) + (\lambda -1)\sum \log(y_i).
\]

```{r}
library(MASS)
library(faraway)

savings_model = lm(sr~., data=savings)
boxcox(savings_model, plotit=TRUE)
```

```{r}
boxcox(savings_model, plotit=TRUE, lambda = seq(0.5, 1.5, by=0.1))
```
```{r}
plot(fitted(savings_model), resid(savings_model), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)

library(lmtest)
bptest(savings_model)
shapiro.test(resid(savings_model))
```

```{r}
gala_model = lm(Species ~ Area + Elevation + Nearest + Scruz + Adjacent, data = gala)
plot(fitted(gala_model), resid(gala_model), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
```

```{r}
boxcox(gala_model, lambda = seq(-0.25, 0.75, by = 0.05), plotit = TRUE)
# lambda = 0.3
```
which suggests a transformation of the form
$$ \frac{y^\lambda - 1}{\lambda} = \frac{y^0.3 - 1}{0.3}$$
```{r}
gala_model_cox = lm((((Species^0.3)-1)/0.3 )~Area + Elevation + Nearest + Scruz + Adjacent, data = gala )

plot(fitted(gala_model_cox), resid(gala_model_cox), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
```

# 2. Predictor Transformation

```{r}
autompg = read.csv("auto-mpg.csv")
colnames(autompg)[-1]<-c("cyl", "disp","hp","wt","acc","year", "origin", "domestic")
autompg$hp<- as.numeric(autompg$hp)
str(autompg)
```

```{r}
par(mfrow= c(1,2))
mpg_hp = lm(mpg~hp, data=autompg)
plot(mpg~hp, data=autompg, col= "dodgerblue", pch = 20, cex= 1.5)
abline(mpg_hp, col = "darkorange", lwd=2)
plot(fitted(mpg_hp), resid(mpg_hp), col="dodgerblue" , pch = 20, cex=1.5,
     xlab= "Fitted", ylab="Residuals")
abline(h=0, lty=2, col="darkorange", lwd= 2)
```

## 2.1 Polynomials

```{r}
marketing = read.csv("marketing.csv")
plot(sales ~ advert, data = marketing, 
     xlab = "Advert Spending (in $100,00)", ylab = "Sales (in $100,00)",
     pch = 20, cex = 2, col="grey")
```
$$ Y_i = \beta_0 + \beta_1 x_i + \beta_2 + x_i ^2 + \epsilon_i$$
```{r}
mark_mod = lm(sales~ advert, data= marketing)
summary(mark_mod)

mark_mod_poly2 = lm(sales~ advert+ I(advert^2), data= marketing)
summary(mark_mod_poly2)
```
```{r}
n = length(marketing$advert)
X = cbind(rep(1, n), marketing$advert, marketing$advert ^ 2)
solve(t(X) %*% X) %*% t(X) %*% marketing$sales

```
```{r}
mark_mod_poly3 = lm(sales ~ advert + I(advert ^ 2) + I(advert ^ 3), data = marketing)
summary(mark_mod_poly3)
```
```{r}
plot(sales ~ advert, data = marketing, 
     xlab = "Advert Spending (in $100,00)", ylab = "Sales (in $100,00)",
     pch = 20, cex = 2)
abline(mark_mod, lty = 2, col = "green", lwd = 2)
xplot = seq(0, 16, by = 0.01)
lines(xplot, predict(mark_mod_poly2, newdata = data.frame(advert = xplot)),
      col = "blue", lwd = 2)
lines(xplot, predict(mark_mod_poly3, newdata = data.frame(advert = xplot)),
      col = "red", lty = 3, lwd = 3)
```

```{r}
library(ggplot2)
ggplot(data= marketing, aes(x= advert, y=sales)) +
  stat_smooth(method = "lm", se = FALSE, color = "green", formula = y ~ x) +
  stat_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ x + I(x ^ 2)) +
  stat_smooth(method = "lm", se = FALSE, color = "red", formula = y ~ x + I(x ^ 2)+ I(x ^ 3)) +
  geom_point(colour = "black", size = 3)
```


```{r}
set.seed(1234)
x = seq(0, 10)
y = 3 + x + 4 * x ^ 2 + rnorm(11, 0, 20)
plot(x, y, ylim = c(-300, 400), cex = 2, pch = 20)
fit = lm(y ~ x + I(x ^ 2))
#summary(fit)
fit_perf = lm(y ~ x + I(x ^ 2) + I(x ^ 3) + I(x ^ 4) + I(x ^ 5) + I(x ^ 6)
               + I(x ^ 7) + I(x ^ 8) + I(x ^ 9) + I(x ^ 10))

xplot = seq(0, 10, by = 0.1)
lines(xplot, predict(fit, newdata = data.frame(x = xplot)),
      col = "dodgerblue", lwd = 2, lty = 1)
lines(xplot, predict(fit_perf, newdata = data.frame(x = xplot)),
      col = "darkorange", lwd = 2, lty = 2)
```

```{r}
econ = read.csv("fuel_econ.csv")
plot_econ_curve = function(model){
  plot(mpg ~ mph, data = econ, xlab = "Speed (Miles per Hour)", 
       ylab = "Fuel Efficiency (Miles per Gallon)", col = "dodgerblue", 
       pch = 20, cex =2)
  xplot = seq(10, 75, by = 0.1)
  lines(xplot, predict(model, newdata = data.frame(mph = xplot)),
        col = "darkorange", lwd = 2, lty = 1)
}

fit1 = lm(mpg ~ mph, data = econ)
par(mfrow=c(1,2))
plot_econ_curve(fit1)
plot(fitted(fit1), resid(fit1), xlab="Fitted", ylab="Residuals", col= "dodgerblue", pch = 20, cex=2)
abline(h=0, col="darkorange", lwd=2)
```

```{r}
fit2 = lm(mpg~mph+I(mph^2), data= econ)
par(mfrow=c(1,2))
plot_econ_curve(fit2)
plot(fitted(fit2), resid(fit2), xlab="Fitted", ylab="Residuals", col= "dodgerblue", pch = 20, cex=2)
abline(h=0, col="darkorange", lwd=2)
```
```{r}
fit3 = lm(mpg~mph +I(mph^2)+I(mph^3), data=econ)
par(mfrow=c(1,2))
plot_econ_curve(fit3)
plot(fitted(fit3), resid(fit3), xlab="Fitted", ylab="Residuals", col= "dodgerblue", pch = 20, cex=2)
abline(h=0, col="darkorange", lwd=2)
```
```{r}
fit4 = lm(mpg ~ mph + I(mph ^ 2) + I(mph ^ 3) + I(mph ^ 4), data = econ)
summary(fit4)
par(mfrow = c(1, 2))
plot_econ_curve(fit4)
plot(fitted(fit4), resid(fit4), xlab = "Fitted", ylab = "Residuals", 
     col = "dodgerblue", pch = 20, cex =2)
  abline(h = 0, col = "darkorange", lwd = 2)
```
```{r}
fit6 = lm(mpg ~ mph + I(mph ^ 2) + I(mph ^ 3) + I(mph ^ 4) + I(mph ^ 5) + I(mph^6), data = econ)
summary(fit6)
par(mfrow = c(1, 2))
plot_econ_curve(fit6)
plot(fitted(fit6), resid(fit6), xlab = "Fitted", ylab = "Residuals", 
     col = "dodgerblue", pch = 20, cex =2)
  abline(h = 0, col = "darkorange", lwd = 2)
```
```{r}
anova(fit4, fit6)
```


























































