---
title: "Chpater 12 Analysis of Variances"
author: "Kwon, YoungJu"
date: "2018년 6월 26일"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(faraway)
plot(coag~diet, data =coagulation, col=2:5)
```

```{r}
coag_aov = aov(coag~diet, data=coagulation)
summary(coag_aov)
```
```{r}
diets = data.frame(diet= unique(coagulation$diet))
data.frame(diets, coag= predict(coag_aov, diets))
```
- Factor vairables
When performing ANOVA, be sure the grouping variable is a factor varaible. If not, your result will be a linear regression
```{r}
set.seed(42)
response = rnorm(15)
group = c(rep(1,5), rep(2,5), rep(3,5))
bad = data.frame(response,group)
summary(aov(response~group, data=bad))
```
```{r}
good = data.frame(response, group = as.factor(group))
summary(aov(response~group, data=good))
```

- Some Simulation
```{r}
library(broom)

sim_anova = function(n=10, mu_a = 0, mu_b = 0, mu_c = 0, mu_d = 0, sigma = 1, stat = TRUE){
  # create data from one-way ANOVA model with four groups of equal size
  
  sim_data = data.frame(
    response = c(rnorm(n=n, mean=mu_a, sd= sigma),
                 rnorm(n=n, mean=mu_b, sd= sigma),
                 rnorm(n=n, mean=mu_c, sd= sigma),
                 rnorm(n=n, mean=mu_d, sd= sigma)),
    group = c(rep("A", times = n), rep("B", times=n),
              rep("C", times=n), rep("D", times = n))
    )
  
  # obtain F-statstic and p-value for testing difference of means
  aov_results = aov(response ~group, data= sim_data)
  f_stat = glance(aov_results)$stat
  p_val = glance(aov_results)$p.val
  
  ifelse(stat, f_stat, p_val)
}

f_stats = replicate(n = 5000, sim_anova(stat = TRUE))
hist(f_stats, breaks = 100, prob = TRUE, border = "dodgerblue", main = "Empirical Distribution of F")
curve(df(x, df1 = 4 - 1, df2 = 40 - 4), col = "darkorange", add = TRUE, lwd = 2)
```

- Power
$$ Power = P(Reject H_{0} | H_{0} False ) $$
Effect size
Noise level
Significance level(alpha)
Sample size
Balanced design

```{r}
p_vals = replicate(n = 1000, sim_anova(mu_a=-1, mu_b = 0, mu_c = 0, mu_d=1
                                       ,sigma = 1.5, stat=FALSE))
mean(p_vals<0.05)
```
```{r}
mean(p_vals<0.01)
```
```{r}
p_vals = replicate(n = 1000, sim_anova(mu_a = -2, mu_b = 0, mu_c = 0, mu_d = 2, 
                                     sigma = 2.0, stat = FALSE))
mean(p_vals < 0.05)
```

- Post Hoc Testing
Test all possible comparisons of two means

```{r}
with(coagulation, pairwise.t.test(coag, diet, p.adj= "none"))
```

- Two-Way Anova

$$ y_{ijk} = \mu +\alpha_{i}+\beta_{j}+(\alpha\beta)_{ij} + \epsilon_{ijk}$$

Rats posison: 3 levels
Rats treat: 4 levels
```{r}
par(mfrow = c(1,2))
with(rats, interaction.plot(poison, treat, time ,lwd =2, col=1:4))
with(rats, interaction.plot(treat, poison, time, lwd=2, col=1:3))
```
interaction.plot function help us visualize the effect of one factor, as we move through the lvels of another factor.

If there is not interaction, we would expect to see parallel lines.
Thus, there can be an effect on the response, but the difference between the levels of the other factor should still be the same.

```{r}
rats_int = aov(time~poison * treat, data=rats) # interaction model
rats_add = aov(time~poison + treat, data= rats) # additive model
rats_pois = aov(time~poison, data=rats) # single factor m
rats_treat = aov(time~treat, data= rats) # single factor m
rats_null = aov(time~1, data= rats) # null model

rats_table = expand.grid(poison= unique(rats$poison),
                         treat = unique(rats$treat))
rats_table
```
```{r}
matrix(paste0(rats_table$poison, "-", 
            rats_table$treat),4,3, byrow=TRUE)
```

```{r}
get_est_means = function(model, table){
  mat = matrix(predict(model, table), nrow =4, ncol=3, byrow=TRUE)
  colnames(mat) = c("I", "II", "III")
  rownames(mat) = c("A", "B","C","D")
  mat
}
```
```{r}
knitr::kable(get_est_means(model=rats_int, table=rats_table))
knitr::kable(get_est_means(model=rats_add, rats_table))
```
```{r}
additive_means = get_est_means(model = rats_add, table = rats_table)
additive_means["A",] - additive_means["B",]

interaction_means = get_est_means(model = rats_int, table = rats_table)
interaction_means["A",] - interaction_means["B",]
```

```{r}
knitr::kable(get_est_means(model = rats_pois, table = rats_table))
knitr::kable(get_est_means(model = rats_treat, table = rats_table))
knitr::kable(get_est_means(model = rats_null, table = rats_table))

```
Model Hierarchy
interaction test
```{r}
print("interaction model")
summary(aov(time ~ poison * treat, data = rats))


print("additive model")
TukeyHSD(aov(time ~ poison + treat, data = rats))
```








