---
title: "STAT420 HW3"
author: "YoungJu Kwon"
date: ''
output:
  html_document: 
    toc: yes
  pdf_document: default
urlcolor: cyan
---

***

## Exercise 1 (Using `lm` for Inference)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Fit the following simple linear regression model in `R`. Use heart weight as the response and body weight as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `cat_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**Solution:**

```{r}
library(MASS)
cat_model= lm(Hwt~Bwt, data=cats)
summary(cat_model)$coefficients[2,3] # t-value
summary(cat_model)$coefficients[2,4] # p-value

```


- $H_0: \beta_1 = 0$
- $H_1: \beta_1 \neq 0$

- Test statistic: $t = `r summary(cat_model)$coefficients["Bwt", "t value"]`$
- P-value: $`r summary(cat_model)$coefficients["Bwt", "Pr(>|t|)"]`$
- Decision: **Reject** $H_0$ at $\alpha = 0.05$.
- Conclusion: There is a linear relationship between heart weight and body weight.

**(b)** Calculate a 90% confidence interval for $\beta_1$. Give an interpretation of the interval in the context of the problem.

**Solution:**

```{r}
confint(cat_model,"Bwt", level=0.90)
```

A 90% confidence interval for $\beta_1$ is given by

\[
(`r confint(cat_model, "Bwt", level = 0.90)[1]`, `r confint(cat_model, "Bwt", level = 0.90)[2]`).
\]

Notice that this interval does **not** contain 0, which suggests that 0 is not a plausible value for $\beta_1$. This notion matches our result from the previous hypothesis test.

Interpretation: We are 90% confident that given a 1-kilogram increase in body weight, the average increase in heart weight will be between `r confint(cat_model, "Bwt", level = 0.90)[1]` and `r confint(cat_model, "Bwt", level = 0.90)[2]` grams.

**(c)** Calculate a 99% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

**Solution:**

```{r, solution = TRUE}
confint(cat_model, "(Intercept)", level=0.99)
```

A 90% confidence interval for $\beta_0$ is given by

\[
(`r confint(cat_model, "(Intercept)", level = 0.99)[1]`, `r confint(cat_model, "(Intercept)", level = 0.99)[2]`).
\]

Interpretation: Mathematically, we are 99% confident that for a body weight of 0 kilograms, the average heart weight will be between `r confint(cat_model, "(Intercept)", level = 0.99)[1]` and `r confint(cat_model, "(Intercept)", level = 0.99)[2]` grams.

However, this confidence interval has no **practical** explanation because it is nonsense to consider the heart weight of a cat that weighs 0 kilograms.

**(d)** Use a 99% confidence interval to estimate the mean heart weight for body weights of 2.1 and 2.8 kilograms. Which of the two intervals is wider? Why?

**Solution:**

```{r, solution = TRUE}
new_Bwt = data.frame(Bwt= c(2.1, 2.8))
heart_weight_ci = predict(cat_model, newdata = new_Bwt, level = 0.99, interval = c("confidence"))
heart_weight_ci[1,3]-heart_weight_ci[1,2]
heart_weight_ci[2,3]-heart_weight_ci[2,2]
```


We are 99% confident that the mean heart weight for a body weight of 2.1 kilograms is in the interval

\[
(`r heart_weight_ci[1, "lwr"]`, `r heart_weight_ci[1, "upr"]`).
\]

We are 99% confident that the mean heart weight for a body weight of 2.8 kilograms is in the interval

\[
(`r heart_weight_ci[2, "lwr"]`, `r heart_weight_ci[2, "upr"]`).
\]

```{r, solution = TRUE}
mean(cats$Bwt)
range(cats$Bwt)
```

The interval for a body weight of 2.1 kilograms is larger since it is further from the sample mean body weight. Also, note that both body weights fall within the range of observed body weights.

```{r, solution = TRUE}
heart_weight_ci = unname(heart_weight_ci) # removes name information for future display
heart_weight_ci[, 2:3]
diff(heart_weight_ci[1, 2:3])
diff(heart_weight_ci[2, 2:3])
diff(heart_weight_ci[1, 2:3]) < diff(heart_weight_ci[2, 2:3])
```

**(e)** Use a 99% prediction interval to predict the heart weight for body weights of 2.8 and 4.2 kilograms.

**Solution:**

```{r, solution = TRUE}
new_Bwt = data.frame(Bwt=c(2.8,4.2))
heart_weight_pi = predict(cat_model, newdata = new_Bwt, 
                          interval = c("prediction"), level = 0.99)
```


We are 99% confident that a *new* **observation** of heart weight for a body weight of 2.8 kilograms is in the interval

\[
(`r heart_weight_pi[1, "lwr"]`, `r heart_weight_pi[1, "upr"]`).
\]

We are 99% confident that a *new* **observation** of heart weight for a body weight of 4.2 kilograms is in the interval

\[
(`r heart_weight_pi[2, "lwr"]`, `r heart_weight_pi[2, "upr"]`).
\]

Note that the prediction interval for a body weight of 2.8 kilograms is wider than the confidence interval for the same body weight found in **(d)**.

```{r, solution = TRUE}
heart_weight_pi = unname(heart_weight_pi)
diff(heart_weight_ci[2, 2:3]) < diff(heart_weight_pi[1, 2:3])
```

**(f)** Create a scatterplot of the data. Add the regression line, 90% confidence bands, and 90% prediction bands.

**Solution:**

```{r, fig.height = 6, fig.width = 8, solution = TRUE}
Bwt_grid = seq(min(cats$Bwt), max(cats$Bwt), by=0.01)
dist_ci_band = predict(cat_model, newdata = data.frame(Bwt=Bwt_grid), 
                       interval = c("confidence"), level=0.9)
dist_pi_band = predict(cat_model, newdata = data.frame(Bwt=Bwt_grid), 
                       interval = c("prediction"), level=0.9)
plot(cats$Hwt~cats$Bwt, xlab="Body Weight", ylab="Heart Weight", cex=2, col="grey", pch=20, ylim=c(min(dist_pi_band), max(dist_pi_band)))
abline(cat_model, lwd = 5, col="darkorange")
lines(Bwt_grid, dist_ci_band[,"lwr"], col="dodgerblue", lwd=3, lty=2)
lines(Bwt_grid, dist_ci_band[,"upr"], col="dodgerblue", lwd=3, lty=2)
lines(Bwt_grid, dist_pi_band[,"lwr"], col="dodgerblue", lwd=3, lty=3)
lines(Bwt_grid, dist_pi_band[,"upr"], col="dodgerblue", lwd=3, lty=3)
points(mean(cats$Bwt), mean(cats$Hwt), pch="+", cex=3)
```

Notice that, while the vast majority of the data points are within the prediction bands, very few points are within the confidence bands.

**(g)** Use a $t$ test to test:

- $H_0: \beta_1 = 4$
- $H_1: \beta_1 \neq 4$

Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**Solution:**

```{r, solution = TRUE}
beta1_hat = coefficients(cat_model)[2]
SE = summary(cat_model)$coefficient[2,2]
t_val = (beta1_hat -4)/SE
p_val = 2*pt(abs(t_val), df=length(resid(cat_model))-2, lower.tail = FALSE)
t_val
p_val
```

- Test statistic: $t = `r t_val`$
- P-value: $`r p_val`$
- Decision: **Fail to reject** $H_0$ at $\alpha = 0.05$.

***

## Exercise 2 (More `lm` for Inference)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will re-perform the data cleaning done in the previous homework.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and wind speed as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_wind_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**Solution:**

```{r, solution = TRUE}
ozone_wind_model = lm(ozone~wind, data= Ozone)
summary(ozone_wind_model)
t_stat = summary(ozone_wind_model)$coefficients["wind", "t value"]
p_val = summary(ozone_wind_model)$coefficients["wind", "Pr(>|t|)"]
```


- $H_0: \beta_1 = 0$, $Y_i = \beta_0 + \epsilon_i$
- $H_1: \beta_1 \neq 0$, $Y_i = \beta_0 + \beta_1 x_i + \epsilon_i$

- Test statistic: $t = `r summary(ozone_wind_model)$coefficients["wind", "t value"]`$
- (Alternative) Test statistic: $F = `r summary(ozone_wind_model)$fstat["value"]`$
- P-value: $`r summary(ozone_wind_model)$coefficients["wind", "Pr(>|t|)"]`$
- Decision: **Fail to Reject** $H_0$ at $\alpha = 0.01$.
- Conclusion: There is **not** a linear relationship between ozone and wind speed.

**(b)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and temperature as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_temp_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**Solution:**

```{r, solution = TRUE}
ozone_temp_model = lm(ozone~temp, data= Ozone)
summary(ozone_temp_model)
t_stat = summary(ozone_temp_model)$coefficients["temp", "t value"]
p_val = summary(ozone_temp_model)$coefficients["temp", "Pr(>|t|)"]
```


- $H_0: \beta_1 = 0$, $Y_i = \beta_0 + \epsilon_i$
- $H_1: \beta_1 \neq 0$, $Y_i = \beta_0 + \beta_1 x_i + \epsilon_i$

- Test statistic: $t = `r summary(ozone_temp_model)$coefficients["temp", "t value"]`$
- (Alternative) Test statistic: $F = `r summary(ozone_temp_model)$fstat["value"]`$
- P-value: $`r summary(ozone_temp_model)$coefficients["temp", "Pr(>|t|)"]`$
- Decision: **Reject** $H_0$ at $\alpha = 0.01$.
- Conclusion: There is a linear relationship between ozone and temperature.

***

## Exercise 3 (Simulating Sampling Distributions)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = -5$
- $\beta_1 = 3.25$
- $\sigma^2 = 16$

We will use samples of size $n = 50$.

**(a)** Simulate this model $2000$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_0$ and $\hat{\beta}_1$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19920502
set.seed(birthday)
n = 50
x = seq(0, 10, length = n)
```

**Solution:**

```{r, solution = TRUE}
num_samples = 2000
n=50 # sample size
beta_0 = -5
beta_1 = 3.25
sigma = 4

x = seq(0,10, length = n)
beta_0_hats = rep(0,num_samples)
beta_1_hats = rep(0,num_samples)
set.seed(19920502)
for(i in 1:num_samples){
  eps = rnorm(n, mean = 0, sd = sigma)
  y = beta_0+beta_1*x+eps
  
  sim_model = lm(y~x)
  beta_0_hats[i] = coefficients(sim_model)[1]
  beta_1_hats[i] = coefficients(sim_model)[2]
}
```

**(b)** Create a table that summarizes the results of the simulations. The table should have two columns, one for $\hat{\beta}_0$ and one for $\hat{\beta}_1$. The table should have four rows:

- A row for the true expected value given the known values of $x$
- A row for the mean of the simulated values
- A row for the true standard deviation given the known values of $x$
- A row for the standard deviation of the simulated values

```{r, solution = TRUE}
Sxx = sum((x - mean(x)) ^ 2)
(var_beta_1_hat = sigma ^ 2 / Sxx)
(var_beta_0_hat = sigma ^ 2 * (1 / num_samples + mean(x) ^ 2 / Sxx))
result_table = data.frame(beta0=c(beta_0, mean(beta_0_hats), sqrt(var_beta_0_hat), sd(beta_0_hats)), beta1 =c(beta_1, mean(beta_1_hats), sqrt(var_beta_1_hat), sd(beta_1_hats)))
result_table
```

| Value         | $\hat{\beta}_0$                             | $\hat{\beta}_1$       |
|---------------|---------------------------------------------|-----------------------|
| $\text{E}[]$  | `r beta_0`                                  | `r beta_1`            |
| `mean()`      | `r mean(beta_0_hats)`                       | `r mean(beta_1_hats)` |
| $\text{SD}[]$ | `r sigma * sqrt(1 / n + mean(x) ^ 2 / Sxx)` | `r sigma / sqrt(Sxx)` |
| `sd()`        | `r sd(beta_0_hats)`                         | `r sd(beta_1_hats)`   |

**(c)** Plot two histograms side-by-side:

- A histogram of your simulated values for $\hat{\beta}_0$. Add the normal curve for the true sampling distribution of $\hat{\beta}_0$.
- A histogram of your simulated values for $\hat{\beta}_1$. Add the normal curve for the true sampling distribution of $\hat{\beta}_1$.

**Solution:**

```{r fig.height = 5, fig.width = 10, solution = TRUE}
# setup plotting
hist(beta_1_hats, prob=TRUE, breaks=20, xlab=expression(hat(beta)[1]), border="dodgerblue")
curve(dnorm(x,mean=beta_1, sd=sqrt(var_beta_1_hat)), col="darkorange", add=TRUE, lwd=3)

hist(beta_0_hats, prob = TRUE, breaks = 25, 
     xlab = expression(hat(beta)[0]), main = "", border = "dodgerblue")
curve(dnorm(x, mean = beta_0, sd = sqrt(var_beta_0_hat)),
      col = "darkorange", add = TRUE, lwd = 3)
```

***

## Exercise 4 (Simulating Confidence Intervals)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 5$
- $\beta_1 = 2$
- $\sigma^2 = 9$

We will use samples of size $n = 25$.

Our goal here is to use simulation to verify that the confidence intervals really do have their stated confidence level. Do **not** use the `confint()` function for this entire exercise.

**(a)** Simulate this model $2500$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_1$ and $s_e$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

**Solution:**

```{r, solution = TRUE}
num_samples = 2500
n=25 # sample size
beta_0 = 5
beta_1 = 2
sigma = 3

set.seed(19920502)
#x = seq(0,10, length = n)
x = seq(-10,10, length = n)
#x = seq(0,2.5, length = n)
beta_0_hats = rep(0,num_samples)
se = rep(0,num_samples)
for(i in 1:num_samples){
  eps = rnorm(n, mean = 0, sd = sigma)
  y = beta_0+beta_1*x+eps
  
  sim_model <- lm(y~x)
  beta_1_hats[i] = coefficients(sim_model)[2]
  se[i] = summary(sim_model)$sigma
}
```


**(b)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 95% confidence interval. Store the lower limits in a vector `lower_95` and the upper limits in a vector `upper_95`. Some hints:

- You will need to use `qt()` to calculate the critical value, which will be the same for each interval.
- Remember that `x` is fixed, so $S_{xx}$ will be the same for each interval.
- You could, but do not need to write a `for` loop. Remember vectorized operations.

**Solution:**

Recall,

\[
\hat{\beta}_1 \pm t_{\alpha/2, n - 2} \cdot \frac{s_e}{\sqrt{S_{xx}}}
\]

```{r, solution = TRUE}
Sxx = sum((x-mean(x))^2)
crit = -qt(0.05/2, df=n-2)
lower_95 = beta_1_hats - crit * se/sqrt(Sxx)
upper_95 = beta_1_hats + crit * se/sqrt(Sxx)
```


**(c)** What proportion of these intervals contains the true value of $\beta_1$?

**Solution:**

```{r, solution = TRUE}
mean(lower_95<2 &2<upper_95)
```

Unsurprisingly, the result is near 95%. (If we increased the number of simulations, this value should move closer to 0.95.)

**(d)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.05$?

**Solution:**

```{r, solution = TRUE}
1-mean(lower_95<0 & 0<upper_95)
```


***

## Exercise 5 (Prediction Intervals "without" `predict`)

Write a function named `calc_pred_int` that performs calculates prediction intervals:

$$
\hat{y}(x) \pm t_{\alpha/2, n - 2} \cdot s_e\sqrt{1 + \frac{1}{n}+\frac{(x-\bar{x})^2}{S_{xx}}}.
$$

for the linear model

$$
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i.
$$

**(a)** Write this function. You may use the `predict()` function, but you may **not** supply a value for the `level` argument of `predict()`. (You can certainly use `predict()` any way you would like in order to check your work.)

The function should take three inputs:

- `model`, a model object that is the result of fitting the SLR model with `lm()`
- `newdata`, a data frame with a single observation (row)
    - This data frame will need to have a variable (column) with the same name as the data used to fit `model`.
- `level`, the level (0.90, 0.95, etc) for the interval with a default value of `0.95`

The function should return a named vector with three elements:

- `estimate`, the midpoint of the interval
- `lower`, the lower bound of the interval
- `upper`, the upper bound of the interval

**Solution:**

```{r, solution = TRUE}
calc_pred_int <-function(model, newdata, level=0.95){
  
  # get t-value
  n = length(resid(model))
  alpha = 1-level
  crit_t = abs(qt(alpha/2, df=n-2))
  
  # get interval from predict()
  crit_t_95 = abs(qt(0.05/2, df=n-2))
  int_level = predict(model, newdata, interval=c("prediction"))
  estimate = int_level[1]
  upr_level = int_level[2]
  lwr_level = int_level[3]
  margin_level = (upr_level-lwr_level)/2
  
  # calculate confidence interval
  SE = margin_level/crit_t_95
  return(c(estimate, estimate+crit_t*SE, estimate-crit_t*SE))
}

newcat_1 = data.frame(Bwt = 4.0)
predict(cat_model, newcat_1, interval = c("prediction"), level=0.95)
calc_pred_int(cat_model, newcat_1)
```

In practice we would never write a function like this, but here it helps us to better understand the construction of the interval.

**(b)** After writing the function, run this code:

```{r, eval = FALSE}
newcat_1 = data.frame(Bwt = 4.0)
calc_pred_int(cat_model, newcat_1)
```

**Solution:**

```{r, solution = TRUE}
newcat_1 = data.frame(Bwt = 4.0)
calc_pred_int(cat_model, newcat_1)
```

**(c)** After writing the function, run this code:

```{r, eval = FALSE}
newcat_2 = data.frame(Bwt = 3.3)
calc_pred_int(cat_model, newcat_2, level = 0.99)
```

**Solution:**

```{r, solution = TRUE}
newcat_2 = data.frame(Bwt = 3.3)
calc_pred_int(cat_model, newcat_2, level = 0.99)
```


